PInSoRo: A Dataset for Deep-learning of Human-Robot Social Interactions

### 2017-06-21 paper.tex
WIP

More recently, Ogata's team~\cite{yang2017repeatable} has demonstrated how an
adequately configured RNN is able to learn a complex robotic task (folding soft
objects like towels using a dual-arm mobile manipulator) from only \emph{35}
demonstrations, one demonstration being a $\approx$ 70 seconds-long teleoperated
sequence. The network inputs are the raw video stream from the head camera and the
12 DoF of the two arms and the total training dataset comprises of about 28000
steps of data.

We argue that the complexity of mechanisms that such neural networks have been
able to uncover and model should invite our community to question its
applicability to human-robot interactions in general, and sustained, natural
child-robot interactions in particular.

However, the current lack of an appropriate HRI dataset suitable for the training of
neural networks hinders this investigation. We present hereafter such a dataset,
and detail our acquisition methodology.

\subsection{Machine learning and social behaviour}

Using interaction datasets to teach robots how to socially behave has been
previously explored, and can be considered as an extension of the traditional
learning from demonstration (LfD) paradigms to social interactions (for
instance~\cite{nehaniv2007imitation,mohammad2015interaction}). Previous examples
have generally focused on low-level recognition or generation of short,
self-standing behaviours, including social gestures~\cite{nagai2005learning} and
gazing behaviours~\cite{calinon2006teaching}.

@@

Kim \etal~\cite{kim2015pororobot} highlight that applying deep learning to
visual scene information in an HRI scenario was successful, but that
generating behaviours for the robot to be able to act in a dynamic and uncertain
environment remains a challenge.

%They rely on unsupervised learning, with a
%training stage consisting in clustering behavioural states (a state consist in
%\emph{behavioural elements} ( joint states of the two interacting agents,
%associated to the current speech elements, as well as the current joint
%\emph{spatial formation}), followed by the {WIP]

These examples show the burgeoning interest of our community for the automatic
learning of social interactions, but also highlight the lack of structure of
these research efforts, as further illustrated by the quasi-absence of public
and large datasets of human-robot interactions. To our best knowledge, only the
$H^3R$ Explanation Corpus~\cite{mohammad2008h} and the Vernissage
Corpus~\cite{jayagopi2013vernissage} have been published to date. The $H^3R$
Explanation Corpus is a human-human and human-robot dataset focusing on a
``assembly/disassembly explanation'' task and includes physiological signals (22
human-robot interactions), but is not publicly available. the Vernissage Corpus
includes one museum guide robot interacting with two people (13 interactions in
total), with recordings and annotations of poses and speech audio (stated to be
publicly available). Both these corpora are however too small for
machine-learning applications.

>>>>

More recently, Ogata's team~\cite{yang2017repeatable} has demonstrated how an
adequately configured RNN is able to learn a complex robotic task (folding soft
objects like towels using a dual-arm mobile manipulator) from only \emph{35}
demonstrations, one demonstration being a $\approx$ 70 seconds-long teleoperated
sequence. The network inputs are the raw video stream from the head camera and the
12 DoF of the two arms and the total training dataset comprises of about 28000
steps of data. Successfully folding towels entails an explicit sequencing of
actions (therefore implicit temporal modeling). The fact that such a complex
process can be successfully learned from a very small dataset certainly opens
many prospects.

Indeed, we argue that the complexity of mechanisms that such neural networks
have been able to uncover and model should invite our community to question its
applicability to human-robot interactions in general, and sustained, natural
child-robot interactions in particular.
However, the current lack of an appropriate HRI dataset suitable for the training of
neural networks hinders this investigation. We present hereafter such a dataset,
and detail our acquisition methodology.

\subsection*{Machine Learning and Social Behaviours}

The use of interaction datasets to teach robots how to socially behave has been
previously explored, and can be considered as the extension of the traditional
learning from demonstration (LfD) paradigms to social interactions (for
instance~\cite{nehaniv2007imitation,mohammad2015interaction}). Previous examples
have generally focused on low-level recognition or generation of short,
self-standing behaviours, including social gestures~\cite{nagai2005learning} and
gazing behaviours~\cite{calinon2006teaching}.

@@

Kim \etal~\cite{kim2015pororobot} highlight that applying deep learning to
visual scene information in an HRI scenario was successful, but that
generating behaviours for the robot to be able to act in a dynamic and uncertain
environment remains a challenge.
They rely on unsupervised learning, with a
training stage consisting in clustering behavioural states (a state consist in
\emph{behavioural elements} (joint states of the two interacting agents,
associated to the current speech elements, as well as the current joint
\emph{spatial formation}), followed by the \todo[inline]{finish that}.

These examples show the burgeoning interest of our community for the automatic
learning of social interactions, but also highlight the lack of structure of
these research efforts, as further illustrated by the lack of large, open
datasets of social interactions, suitable for machine-learning applications.

One of the most interesting datasets is the \emph{Multimodal Dyadic Behavior
Dataset} (\emph{MMDB},~\cite{rehg2013decoding}). It comprises of 160 sessions of
3 to 5 minute child-adult interactions. During these interactions, the
experimenter plays with toddlers (1.5 to 2.5 years old) in semi-structured ways
which was designed to
elicit a broad range of social behaviors
